Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-05-03 paper/graph.tex
More edits; consistent notation.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@602 3fc257bb-8550-0410-81b3-a5c103da83f7

We begin with a MapReduce procedure for creating large, sparse,
randomized graphs, since they are the input for the algorithms
discussed below.  R-MAT graphs ~\cite{RMAT} are recursively generated
graphs with power-law degree distributions.  They are commonly used to
represent web and social networks.  The user specifices 6 parameters
which define the graph: the number of vertices $N$ and edges $M$, and
4 parameters $a$, $b$, $c$, $d$ which sum to 1.0 and are discussed
below.  The algorithm in Figure \ref{fig:rmat} generates $M$ unique
non-zero entries in a sparse $NxN$ matrix $A$, where each entry
$A_{ij}$ represents an edge between graph vertices $(V_i,V_j)$.

@@

$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> 1 Map: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> 1 Collate \\
\> 1 Reduce: \> Remove duplicate edges \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$

@@

In the map() operation, each of $P$ processors generates a $1/P$
fraction of the desired edges.  A single random $i,j$ edge is computed
recursively as follows.  Pick a random quadrant of the $A$ matrix with
relative probabilities $a$, $b$, $c$, and $d$.  Treat the chosen
quadrant as a sub-matrix and select a random quadrant within it, in
the same manner.  Repeat this process $n$ times where $N = 2^n$.  At
the end of the recursion, the final ``quadrant'' is non-zero matrix
element $A_{ij}$.

The map() will often generate some small number of duplicate edges.
The collate() and reduce() operations remove the duplicates.  The
entire map-collate-reduce sequence is repeated until the number of
resulting key/value pairs $N_{kv} = M$.  For reasonably sparse graphs
this typically takes only a few iterations.

Note that the degree distribution of vertices in the graph depends on
the choice of parameters $a$, $b$, $c$, $d$.  If one of the four
values is larger than the other 3, a highly skewed distribution
results.  Variants of the above algorithm can be used when $N$ is not
a power-of-two, to generate graphs with weighted edges (assign a
numeric value to the $A_{ij}$ edge), graphs without self edges
(require $i != j$), or graphs with undirected edges (require $i < $j).
Or the general R-MAT matrix can be further processed by MapReduce
operations to meet these requirements.

By using more than one MapReduce object, we can improve the performance
of the R-MAT generation algorithm.  In the improved algorithm
(Figure~\ref{fig:rmat2}, we emit newly created edges into MapReduce object
$E_{new}$ and aggregate the edges to processors.  We then add those edges
to the MapReduce object $E_{old}$, containing all previously generated
unique edges.  We convert the total set of key/values to key/multivalues
and use a reduce operation to remove duplicates.  Performance improvements
in this algorithm result from the reduced number of key/values in the aggregate
operation, compared to the collate operation in the original algorithm, 
particularly in later iterations of the generation where few
edges are created.

@@

$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> 1 Map $E_{new}$: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> 1 Aggregate $E_{new}$. \\
\> 1 Add $E_{new}$ to $E_{old}$. \\
\> 1 Compress $E_{old}$: \> Remove duplicate edges \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$

@@

The PageRank algorithm assings a relative numeric rank to each
vertex in a graph.

It models the web as a directed graph $G(V,E)$, with each vertex $v_i
\in V$ representing a web page and each edge $E_{ij} \in E$
representing a hyperlink from $v_i$ TO $v_j$.  The probability of
moving from $v_i$ to another vertex $v_j$ is $\alpha/d_{out}(v_i) +
(1-\alpha)/|V|$, where $\alpha$ is a user-defined parameter (usually
0.8-0.9), $d_{out}(v_i)$ is the outdegree of vertex $v_i$, and $|V|$ is
the cardinality of $V$.  The first term represents the probability of
following a given link on page $v_i$; the second represents the
probability of moving to a random page.  For pages with no outlinks,
the first term is $\alpha/|V|$, indicating equal likelihood to move to
any other page.  Equivalently, the graph can be represented by a
matrix $A$~\cite{LangvilleMeyer05a}, with matrix entries $A_{ij} =
\alpha/d_{out}(v_i)$ if vertex $v_i$ links to $v_j$.  The PageRank
algorithm, then, is essentially a power-method iteration in which the
dominating computation is matrix-vector multiplication $A^T x=y$,
where $x$ is the PageRank vector from the previous iteration.

The algorithm in Figure \ref{fig:pr} performs these
iterations.

@@

%Input from R-MAT generator: \\
%\> MapReduce object $V$ containing all graph vertices:  Key=$v_i$, Value = NULL.  \\
%\> MapReduce object $E$ containing all graph edges: Key=$v_i$, Value = $v_j$. \\
%Initialize MapReduce object for matrix $A$  and \\
%MapReduce object $MT$ containing all vertices with no outlinks.\\
%\> Copy: $A = E$ \\
%\> Add: $V$ to $A$ \\
%\> Collate $A$: Vertex is Key \\
%\> Reduce $A$:  \\
%\> \> Input:  Key = $v_i$; MultiValue = NULL,$v_j$,$v_k$,... \\
%\> \> If (${nvalues} > 1$) \\
%\> \> \> Compute: $d_{out}(v_i) = {nvalues} - 1$. \\
%\> \> \> Output:  For each $v_j$ in MultiValue, Key = $v_i$; Value = ($v_j$, $\alpha/d_{out}(v_i)$). \\
%\> \> else \\
%\> \> \> Output to $MT$: Key = $v_i$; Value = 0. \\
%Initialize MapReduce object for pagerank vector $x$. \\
%\> Map:   \\
%\> \> Input:  Graph vertices $V$. \\
%\> \> Output:  For each vertex, Key = $v_i$; Value = $1/|V|$. \\
Input: \\
\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, a_{ij}]$ \\
\> MapReduce object $MT$ containing indices of empty rows of $A$:  Key = $i$, Value = 0. \\
\> MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$. \\
While (${residual} > {tolerance}$) \\
\> Compute contribution for random jumps and zero-outdegree vertices. \\
\> \> Add: $x$ to $MT$. \\
\> \> Collate $MT$:  Vertex is Key. \\
\> \> Reduce $MT$:   \\
\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$. \\
\> \> \> Compute:  Local contributions $cont$. \\
\> \> \> Output:  If ${nvalues} = 2$, Key = $v_i$; Value = 0. \\
\> \> MPI\_Allreduce with MPI\_SUM to compute global contribution $gcont$. \\
\> Compute $y = A^T x$.  \\
\> \> Copy $x$ to $y$. \\
\> \> Add $A$ to $y$. \\
\> \> Collate $y$:  Key is row index of $A$. \\
\> \> Reduce $y$: \\
\> \> \>  Input: Key = $i$; Multivalue = $i, [j, a_{ij}]$ for nonzeros $a_{ij}$ in $A$ \\
\> \> \>  Output:  Key = $j$; Value = $a_{ij} y_i$. \\
\> \> Collate $y$:  Key is column index of $A$. \\
\> \> Reduce $y$:  \\
\> \> \> Input:  Key = $j$; Multivalue = $a_{ij} y_i$ for all non-zero entries $a_{ij}$ of column $j$ \\
\> \> \> Output:  Key = $j$; Value = $\sum_i a_{ij} y_i$. \\
\> Add contribution $gcont$ to $y$; compute local max norm $norm$ of $y$. \\
\> \> Map $y$:  \\
\> \> \> Input:  Key = $i$, Value = $y_i$. \\
\> \> \> Compute: $y_i = y_i + gcont$; if $y_i > {norm}$, $norm = y_i$. \\
\> \> \> Output:  Key = $i$, Value = $y_i$. \\
\> \> MPI\_Allreduce $norm$ with MPI\_MAX to compute global max norm $gnorm$. \\
\> Scale $y$ by $gnorm$. \\
\> \> Map $y$: \\
\> \> \> Input:  Key = $i$, Value = $y_i$. \\
\> \> \> Output:  Key = $i$, Value = $y_i / {gnorm}$. \\
\> Compute residual. \\
\> \> Add $y$ to $x$. \\
\> \> Collate $x$:  Key is index $i$. \\
\> \> Reduce $x$:   \\
\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$. \\
\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$. \\
\> $x = y$; \\
Output:  PageRank vector $x$.
  \end{tabbing}
 \end{minipage}}\end{center}

@@

Input: \\
$*$\> Pre-aggregated MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, a_{ij}]$ \\
$*$\> Pre-aggregated MapReduce object $MT$ containing indices of empty rows of $A$:  Key = $i$, Value = 0. \\
$*$\> Pre-aggregated MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$. \\
While (${residual} > {tolerance}$) \\
\> Compute contribution for random jumps and zero-outdegree vertices. \\
\> \> Add: $x$ to $MT$. \\
$*$\> \> Convert $MT$:  Vertex is Key. \\
\> \> Reduce $MT$:   \\
\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$. \\
\> \> \> Compute:  Local contributions $cont$. \\
\> \> \> Output:  If ${nvalues} = 2$, Key = $v_i$; Value = 0. \\
\> \> MPI\_Allreduce with MPI\_SUM to compute global contribution $gcont$. \\
\> Compute $y = A^T x$.  \\
\> \> Copy $x$ to $y$. \\
\> \> Add $A$ to $y$. \\
$*$\> \> Convert $y$:  Key is row index of $A$. \\
\> \> Reduce $y$: \\
\> \> \>  Input: Key = $i$; Multivalue = $i, [j, a_{ij}]$ for nonzeros $a_{ij}$ in $A$ \\
\> \> \>  Output:  Key = $j$; Value = $a_{ij} y_i$. \\
\> \> Collate $y$:  Key is column index of $A$. \\
\> \> Reduce $y$:  \\
\> \> \> Input:  Key = $j$; Multivalue = $a_{ij} y_i$ for all non-zero entries $a_{ij}$ of column $j$ \\
\> \> \> Output:  Key = $j$; Value = $\sum_i a_{ij} y_i$. \\
\> Add contribution $gcont$ to $y$; compute local max norm $norm$ of $y$. \\
\> \> Map $y$:  \\
\> \> \> Input:  Key = $i$, Value = $y_i$. \\
\> \> \> Compute: $y_i = y_i + gcont$; if $y_i > {norm}$, $norm = y_i$. \\
\> \> \> Output:  Key = $i$, Value = $y_i$. \\
\> \> MPI\_Allreduce $norm$ with MPI\_MAX to compute global max norm $gnorm$. \\
\> Scale $y$ by $gnorm$. \\
\> \> Map $y$: \\
\> \> \> Input:  Key = $i$, Value = $y_i$. \\
\> \> \> Output:  Key = $i$, Value = $y_i / {gnorm}$. \\
\> Compute residual. \\
\> \> Add $y$ to $x$. \\
$*$\> \> Convert $x$:  Key is index $i$. \\
\> \> Reduce $x$:   \\
\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$. \\
\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$. \\
\> $x = y$; \\
Output:  PageRank vector $x$.
  \end{tabbing}
 \end{minipage}}\end{center}

@@

In MapReduce, global information about edge adjacencies and labeled vertices
is not available.  Instead, following the Bellman-Ford-style 
algorithm of~\cite{SSSPMapReduce, BellmanFord},
every edge of the graph is visited in each iteration,
although labels for an edge's vertices are updated only if the breadth-first
traversal has reached those vertices.  The algorithm begins with a {\it map()}
function loading edges $(v_i, v_j)$ and source vertex $v_s$ with distance 0.
In each iteration, the edges $(v_i, v_j)$ are collated with respect to
$v_i$, and a {\it reduce()} operation updates each $v_j$'s distance if
$v_i$'s distance is non-zero.  The algorithm is fully described in
Figure~\ref{fig:sssp}.

@@

Input:  \\
\> Source vertex $v_s$.\\
\> MapReduce object $V$ containing graph vertices $v_i$ with distances $d_i$ from $v_s$ (initially $\infty$):  \\
\> \> Key = $v_i$; Value = $\infty$.\\
\> MapReduce object $E$ containing graph edges ($v_i, v_j$) with weights $w_{ij}$: \\
\> \>  Key = $v_i$; Value = $[v_j, w_{ij}]$ \\
Map $V$:  Set distance for source vertex $v_s$ to zero. \\
\> Input:  Key = $v_i$; Value = $\infty$ \\
\> Output:  Key = $v_i$; Value = 0 if $v_i = v_s$, otherwise Value = $\infty$. \\
while not done: \\
\> done = 1; \\
\> Add $E$ to $V$. \\
\> Collate $V$:  Key is vertex $v_i$. \\
\> Reduce $V$:  Loop over all edges to perform breadth-first search from labeled vertices. \\
\> \> Input:  Key = $v_i$; Multivalue = edges $[v_j, w_{ij}]$, candidate distances $d_i$. \\
\> \> Compute:  Find smallest candidate distance $d_{min}$ for $v_i$.  \\
\> \> \> If $d_{min} \ne d_i$, done = 0; \\
\> \> Output:  Key = $v_i$; Value = $d_{min}$ (updated distance for $v_i$) \\
\> \> \> and if $d_{min} \ne d_i$, Key = $v_j$; Value = $d_{min} + w_{ij}$ for each edge ($v_i, v_j$). \\
\> MPI\_Allreduce on done with MPI\_MIN. \\
Output: MapReduce object $V$ with Key = $v_i$, Value = $d_i$.

@@

Input:  \\
\> Source vertex $v_s$.\\
\> Pre-aggregated MapReduce object $V$ containing graph vertices $v_i$ with distances $d_i$ from $v_s$:  \\
\> \> Key = $v_i$; Value = $\infty$.\\
\> Pre-aggregated MapReduce object $E$ containing graph edges ($v_i, v_j$) with weights $w_{ij}$: \\
\> \>  Key = $v_i$; Value = $[v_j, w_{ij}]$ \\
$U$ = new MapReduce object to store updates to distances. \\
Map $U$:  Add $v_s$ with distance $d_s = 0$ to $U$. \\
\> Input:  $v_s$ \\
\> Output:  Key = $v_s$; Value = 0. \\
while not done: \\
\> done = 1; \\
\> Aggregate $U$. \\
\> Add $U$ to $V$. Empty $U$.\\
\> Compress $V$:  Pick best candidate distance for each vertex. \\
\> \> Input:  Key = $v_i$; Multivalue = candidate distances $d_i$. \\
\> \> Compute:  Find smallest candidate distance $d_{min}$ for $v_i$.  \\
\> \> \> If $d_{min} \ne d_i$, done = 0; \\
\> \> Output to $V$:  Key = $v_i$; Value = $d_{min}$. \\
\> \> Output to $U$: if $d_{min} \ne d_i$, Key = $v_i$; Value = $d_{min}$. \\
\> MPI\_Allreduce on done with MPI\_MIN. \\
\> if not done \\
\> \> Add $U$ to $E$.  Empty $U$. \\
\> \> Compress $E$:  Generate candidate distances for neighbors of changed vertices. \\
\> \> \> Input:  Key = $v_i$; Multivalue = $[v_j, w_{ij}]$ for edges ($v_i, v_j$) \\
\> \> \> \> and distance $d_i$ if updated in previous step. \\
\> \> \> Output to $U$:  If updated distance $d_i$ exists, \\
\> \> \> \> Key = $v_j$; Value = $d_i+w_{ij}$ for each edge ($v_i, v_j$). \\
Output: MapReduce object $V$ with Key = $v_i$, Value = $d_i$.

>>>>

We begin with a MapReduce procedure for creating large, sparse,
randomized graphs, since they are the input for the algorithms
discussed below.  R-MAT graphs ~\cite{RMAT} are recursively generated
graphs with power-law degree distributions.  They are commonly used to
represent web and social networks.  The user specifices six parameters
that define the graph: the number of vertices $N$ and edges $M$, and
four parameters $a$, $b$, $c$, $d$ that sum to 1.0 and are discussed
below.  The algorithm in Figure \ref{fig:rmat} generates $M$ unique
non-zero entries in a sparse $N \times N$ matrix $A$, where each entry
$A_{ij}$ represents an edge between graph vertices $(V_i,V_j)$.

@@

$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> Map: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
           \> \> output Key = $(i,j)$, Value = NULL \\
\> Collate \\
\> Reduce: \> Remove duplicate edges \\
           \> \> input Key = $(i,j)$, MultiValue = one or more NULLs \\
           \> \> output Key = $(i,j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$

@@

In the {\it map()} operation, each of $P$ processors generates a $1/P$
fraction of the desired edges.  A single random edge ($i,j$) is computed
recursively as follows.  Pick a random quadrant of the $A$ matrix with
relative probabilities $a$, $b$, $c$, and $d$.  Treat the chosen
quadrant as a sub-matrix and select a random quadrant within it, in
the same manner.  Repeat this process $n$ times where $N = 2^n$.  At
the end of the recursion, the final ``quadrant'' is non-zero matrix
element $A_{ij}$.

The {\it map()} will often generate some small number of duplicate edges.
The {\it collate()} and {\it reduce()} operations remove the duplicates.  The
entire map-collate-reduce sequence is repeated until the number of
resulting key/value pairs $N_{kv} = M$.  For reasonably sparse graphs
this typically takes only a few iterations.

Note that the degree distribution of vertices in the graph depends on
the choice of parameters $a$, $b$, $c$, $d$.  If one of the four
values is larger than the other three, a highly skewed distribution
results.  Variants of the above algorithm can be used when $N$ is not
a power-of-two, to generate graphs with weighted edges (assign a
numeric value to the $A_{ij}$ edge), graphs without self edges
(require $i \ne j$), or graphs with undirected edges (require $i < j$).
Or the general R-MAT matrix can be further processed by MapReduce
operations to meet these requirements.

By using more than one MapReduce object, we can improve the performance
of the R-MAT generation algorithm.  In the improved algorithm
(Figure~\ref{fig:rmat2}), we emit newly created edges into MapReduce object
$E_{new}$ and {\it aggregate()} the edges to processors.  
We then add those edges
to the MapReduce object $E_{old}$, containing all previously generated
unique edges; the edges of $E_{old}$ are already aggregated to processors
using the same mapping as $E_{new}$.  We can then use the {\it compress()}
function to reduce duplicates locally on each processor.
Performance improvements
in this algorithm result from the reduced number of KV pairs communicated in
{\it aggregate()},
compared to the collate operation in the original algorithm, 
particularly in later iterations of the generation where few
edges are created.

@@

$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> Map $E_{new}$: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
           \> \> output Key = $(i,j)$, Value = NULL \\
\> Aggregate $E_{new}$. \\
\> Add $E_{new}$ to $E_{old}$. \\
\> Compress $E_{old}$: Remove duplicate edges \\
            \> \> input Key = $(i,j)$, MultiValue = one or more NULLs \\
            \> \> output Key = $(i,j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$

@@

The PageRank algorithm assings a relative numeric rank to each
vertex in a graph.
It models the web as a directed graph $G(V,E)$, with each vertex $V_i
\in V$ representing a web page and each edge $(V_i, V_j) \in E$
representing a hyperlink from $V_i$ to $V_j$.  The probability of
moving from $V_i$ to another vertex $V_j$ is $\alpha/d_{out}(V_i) +
(1-\alpha)/|V|$, where $\alpha$ is a user-defined parameter (usually
0.8-0.9), $d_{out}(V_i)$ is the outdegree of vertex $V_i$, and $|V|$ is
the cardinality of $V$.  The first term represents the probability of
following a given link on page $V_i$; the second represents the
probability of moving to a random page.  For pages with no outlinks,
the first term is $\alpha/|V|$, indicating equal likelihood to move to
any other page.  Equivalently, the graph can be represented by a
matrix $A$~\cite{LangvilleMeyer05a}, with matrix entries $A_{ij} =
\alpha/d_{out}(V_i)$ if vertex $V_i$ links to $V_j$.  The PageRank
algorithm, then, is essentially a power-method iteration in which the
dominating computation is matrix-vector multiplication $A^T x=y$,
where $x$ is the PageRank vector from the previous iteration.
The algorithm in Figure \ref{fig:pr} performs these
iterations.

@@

%Input from R-MAT generator: \\
%\> MapReduce object $V$ containing all graph vertices:  Key=$V_i$, Value = NULL.  \\
%\> MapReduce object $E$ containing all graph edges: Key=$V_i$, Value = $V_j$. \\
%Initialize MapReduce object for matrix $A$  and \\
%MapReduce object $MT$ containing all vertices with no outlinks.\\
%\> Copy: $A = E$ \\
%\> Add: $V$ to $A$ \\
%\> Collate $A$: Vertex is Key \\
%\> Reduce $A$:  \\
%\> \> Input:  Key = $V_i$; MultiValue = NULL,$V_j$,$V_k$,... \\
%\> \> If (${nvalues} > 1$) \\
%\> \> \> Compute: $d_{out}(V_i) = {nvalues} - 1$. \\
%\> \> \> Output:  For each $V_j$ in MultiValue, Key = $V_i$; Value = ($V_j$, $\alpha/d_{out}(V_i)$). \\
%\> \> else \\
%\> \> \> Output to $MT$: Key = $V_i$; Value = 0. \\
%Initialize MapReduce object for pagerank vector $x$. \\
%\> Map:   \\
%\> \> Input:  Graph vertices $V$. \\
%\> \> Output:  For each vertex, Key = $V_i$; Value = $1/|V|$. \\
Input: \\
\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, a_{ij}]$ \\
\> MapReduce object $MT$ containing indices of empty rows of $A$:  Key = $i$, Value = 0. \\
\> MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$. \\
While (${residual} > {tolerance}$) \\
\> Compute contribution for random jumps and zero-outdegree vertices. \\
\> \> Add: $x$ to $MT$. \\
\> \> Collate $MT$:  Vertex is Key. \\
\> \> Reduce $MT$:   \\
\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$. \\
\> \> \> Compute:  Local contributions $cont$. \\
\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0. \\
\> \> MPI\_Allreduce with MPI\_SUM to compute global contribution $gcont$. \\
\> Compute $y = A^T x$.  \\
\> \> Copy $x$ to $y$. \\
\> \> Add $A$ to $y$. \\
\> \> Collate $y$:  Key is row index of $A$. \\
\> \> Reduce $y$: \\
\> \> \>  Input: Key = $i$; Multivalue = $i, [j, a_{ij}]$ for nonzeros $a_{ij}$ in $A$ \\
\> \> \>  Output:  Key = $j$; Value = $a_{ij} y_i$. \\
\> \> Collate $y$:  Key is column index of $A$. \\
\> \> Reduce $y$:  \\
\> \> \> Input:  Key = $j$; Multivalue = $a_{ij} y_i$ for all non-zero entries $a_{ij}$ of column $j$ \\
\> \> \> Output:  Key = $j$; Value = $\sum_i a_{ij} y_i$. \\
\> Add contribution $gcont$ to $y$; compute local max norm $norm$ of $y$. \\
\> \> Map $y$:  \\
\> \> \> Input:  Key = $i$, Value = $y_i$. \\
\> \> \> Compute: $y_i = y_i + gcont$; if $y_i > {norm}$, $norm = y_i$. \\
\> \> \> Output:  Key = $i$, Value = $y_i$. \\
\> \> MPI\_Allreduce $norm$ with MPI\_MAX to compute global max norm $gnorm$. \\
\> Scale $y$ by $gnorm$. \\
\> \> Map $y$: \\
\> \> \> Input:  Key = $i$, Value = $y_i$. \\
\> \> \> Output:  Key = $i$, Value = $y_i / {gnorm}$. \\
\> Compute residual. \\
\> \> Add $y$ to $x$. \\
\> \> Collate $x$:  Key is index $i$. \\
\> \> Reduce $x$:   \\
\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$. \\
\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$. \\
\> $x = y$; \\
Output:  PageRank vector $x$.
  \end{tabbing}
 \end{minipage}}\end{center}

@@

Input: \\
$*$\> Pre-aggregated MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, a_{ij}]$ \\
$*$\> Pre-aggregated MapReduce object $MT$ containing indices of empty rows of $A$:  Key = $i$, Value = 0. \\
$*$\> Pre-aggregated MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$. \\
While (${residual} > {tolerance}$) \\
\> Compute contribution for random jumps and zero-outdegree vertices. \\
\> \> Add: $x$ to $MT$. \\
$*$\> \> Convert $MT$:  Vertex is Key. \\
\> \> Reduce $MT$:   \\
\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$. \\
\> \> \> Compute:  Local contributions $cont$. \\
\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0. \\
\> \> MPI\_Allreduce with MPI\_SUM to compute global contribution $gcont$. \\
\> Compute $y = A^T x$.  \\
\> \> Copy $x$ to $y$. \\
\> \> Add $A$ to $y$. \\
$*$\> \> Convert $y$:  Key is row index of $A$. \\
\> \> Reduce $y$: \\
\> \> \>  Input: Key = $i$; Multivalue = $i, [j, a_{ij}]$ for nonzeros $a_{ij}$ in $A$ \\
\> \> \>  Output:  Key = $j$; Value = $a_{ij} y_i$. \\
\> \> Collate $y$:  Key is column index of $A$. \\
\> \> Reduce $y$:  \\
\> \> \> Input:  Key = $j$; Multivalue = $a_{ij} y_i$ for all non-zero entries $a_{ij}$ of column $j$ \\
\> \> \> Output:  Key = $j$; Value = $\sum_i a_{ij} y_i$. \\
\> Add contribution $gcont$ to $y$; compute local max norm $norm$ of $y$. \\
\> \> Map $y$:  \\
\> \> \> Input:  Key = $i$, Value = $y_i$. \\
\> \> \> Compute: $y_i = y_i + gcont$; if $y_i > {norm}$, $norm = y_i$. \\
\> \> \> Output:  Key = $i$, Value = $y_i$. \\
\> \> MPI\_Allreduce $norm$ with MPI\_MAX to compute global max norm $gnorm$. \\
\> Scale $y$ by $gnorm$. \\
\> \> Map $y$: \\
\> \> \> Input:  Key = $i$, Value = $y_i$. \\
\> \> \> Output:  Key = $i$, Value = $y_i / {gnorm}$. \\
\> Compute residual. \\
\> \> Add $y$ to $x$. \\
$*$\> \> Convert $x$:  Key is index $i$. \\
\> \> Reduce $x$:   \\
\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$. \\
\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$. \\
\> $x = y$; \\
Output:  PageRank vector $x$.
  \end{tabbing}
 \end{minipage}}\end{center}

@@

In MapReduce, global information about edge adjacencies and labeled vertices
is not available.  Instead, following the Bellman-Ford-style 
algorithm of~\cite{SSSPMapReduce, BellmanFord},
every edge of the graph is visited in each iteration,
although labels for an edge's vertices are updated only if the breadth-first
traversal has reached those vertices.  The algorithm begins with a {\it map()}
function loading edges $(V_i, V_j)$ and source vertex $V_s$ with distance 0.
In each iteration, the edges $(V_i, V_j)$ are collated with respect to
$V_i$, and a {\it reduce()} operation updates each $V_j$'s distance if
$V_i$'s distance is non-zero.  The algorithm is fully described in
Figure~\ref{fig:sssp}.

@@

Input:  \\
\> Source vertex $V_s$.\\
\> MapReduce object $V$ containing graph vertices $V_i$ with distances $d_i$ from $V_s$ (initially $\infty$):  \\
\> \> Key = $V_i$; Value = $\infty$.\\
\> MapReduce object $E$ containing graph edges ($V_i, V_j$) with weights $w_{ij}$: \\
\> \>  Key = $V_i$; Value = $[V_j, w_{ij}]$ \\
Map $V$:  Set distance for source vertex $V_s$ to zero. \\
\> Input:  Key = $V_i$; Value = $\infty$ \\
\> Output:  Key = $V_i$; Value = 0 if $V_i = V_s$, otherwise Value = $\infty$. \\
while not done: \\
\> done = 1; \\
\> Add $E$ to $V$. \\
\> Collate $V$:  Key is vertex $V_i$. \\
\> Reduce $V$:  Loop over all edges to perform breadth-first search from labeled vertices. \\
\> \> Input:  Key = $V_i$; Multivalue = edges $[V_j, w_{ij}]$, candidate distances $d_i$. \\
\> \> Compute:  Find smallest candidate distance $d_{min}$ for $V_i$.  \\
\> \> \> If $d_{min} \ne d_i$, done = 0; \\
\> \> Output:  Key = $V_i$; Value = $d_{min}$ (updated distance for $V_i$) \\
\> \> \> and if $d_{min} \ne d_i$, Key = $V_j$; Value = $d_{min} + w_{ij}$ for each edge ($V_i, V_j$). \\
\> MPI\_Allreduce on done with MPI\_MIN. \\
Output: MapReduce object $V$ with Key = $V_i$, Value = $d_i$.

@@

Input:  \\
\> Source vertex $V_s$.\\
\> Pre-aggregated MapReduce object $V$ containing graph vertices $V_i$ with distances $d_i$ from $V_s$:  \\
\> \> Key = $V_i$; Value = $\infty$.\\
\> Pre-aggregated MapReduce object $E$ containing graph edges ($V_i, V_j$) with weights $w_{ij}$: \\
\> \>  Key = $V_i$; Value = $[V_j, w_{ij}]$ \\
$U$ = new MapReduce object to store updates to distances. \\
Map $U$:  Add $V_s$ with distance $d_s = 0$ to $U$. \\
\> Input:  $V_s$ \\
\> Output:  Key = $V_s$; Value = 0. \\
while not done: \\
\> done = 1; \\
\> Aggregate $U$. \\
\> Add $U$ to $V$. Empty $U$.\\
\> Compress $V$:  Pick best candidate distance for each vertex. \\
\> \> Input:  Key = $V_i$; Multivalue = candidate distances $d_i$. \\
\> \> Compute:  Find smallest candidate distance $d_{min}$ for $V_i$.  \\
\> \> \> If $d_{min} \ne d_i$, done = 0; \\
\> \> Output to $V$:  Key = $V_i$; Value = $d_{min}$. \\
\> \> Output to $U$: if $d_{min} \ne d_i$, Key = $V_i$; Value = $d_{min}$. \\
\> MPI\_Allreduce on done with MPI\_MIN. \\
\> if not done \\
\> \> Add $U$ to $E$.  Empty $U$. \\
\> \> Compress $E$:  Generate candidate distances for neighbors of changed vertices. \\
\> \> \> Input:  Key = $V_i$; Multivalue = $[V_j, w_{ij}]$ for edges ($V_i, V_j$) \\
\> \> \> \> and distance $d_i$ if updated in previous step. \\
\> \> \> Output to $U$:  If updated distance $d_i$ exists, \\
\> \> \> \> Key = $V_j$; Value = $d_i+w_{ij}$ for each edge ($V_i, V_j$). \\
Output: MapReduce object $V$ with Key = $V_i$, Value = $d_i$.
